---
title: 'Project #1'
author: "Tejaswi Neelapu"
output: html_document
date: "2023-10-29"
---
```{r}
library(nycflights13)
library(tidyverse)

United_Airlines <- flights %>% filter(carrier == "UA") %>% inner_join(weather, by = c("year", "month", "day", "hour", "origin"))
glimpse(United_Airlines)

```


1. Time of day

```{r}
Time_of_day <- United_Airlines %>% select(dep_time, dep_delay) %>%
  mutate(time_of_day = case_when(
    dep_time >= 600 & dep_time < 1200 ~ "Morning",
    dep_time >= 1200 & dep_time < 1800 ~ "Afternoon",
    dep_time >= 1800 & dep_time < 2100 ~ "Evening",
    TRUE ~ "Night"))

Time_of_day
ggplot(Time_of_day, aes(x = time_of_day, y = dep_delay)) +
  geom_boxplot(color = "cadetblue") +
  labs(
    x = "Time of Day",
    y = "Departure Delay",
    title = "Time of Day versus Departure Delays for United Airlines (UA)"
  )
summary(Time_of_day)
```

Comparing by dividing the day into four categories:

Permutation test between Morning and Afternoon:

```{r}
Time_of_day <- United_Airlines %>% select(dep_time, dep_delay) %>%
  mutate(time_of_day = case_when(
    dep_time >= 600 & dep_time < 1200 ~ "Morning",
    dep_time >= 1200 & dep_time < 1800 ~ "Afternoon"))

Time_of_day <- na.omit(Time_of_day)
Time_of_day

observed <- mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Morning"]) - mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Afternoon"])
observed

N <- 10^4-1

sample.size = nrow(Time_of_day)
group.1.size = nrow(Time_of_day[Time_of_day$time_of_day == "Morning",])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Time_of_day$dep_delay[index]) - median(Time_of_day$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```

Permutation test between Afternoon and Evening:

```{r}
Time_of_day <- United_Airlines %>% select(dep_time, dep_delay) %>%
  mutate(time_of_day = case_when(
    dep_time >= 1200 & dep_time < 1800 ~ "Afternoon",
    dep_time >= 1800 & dep_time < 2100 ~ "Evening"))

Time_of_day <- na.omit(Time_of_day)
Time_of_day

observed <- mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Afternoon"]) - mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Evening"])
observed

N <- 10^4-1

sample.size = nrow(Time_of_day)
group.1.size = nrow(Time_of_day[Time_of_day$time_of_day == "Afternoon",])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Time_of_day$dep_delay[index]) - median(Time_of_day$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```

Permutation test between Evening and Night:

```{r}
Time_of_day <- United_Airlines %>% select(dep_time, dep_delay) %>%
  mutate(time_of_day = case_when(
    dep_time >= 600 & dep_time < 1200 ~ "Morning",
    dep_time >= 1200 & dep_time < 1800 ~ "Afternoon",
    dep_time >= 1800 & dep_time < 2100 ~ "Evening",
    TRUE ~ "Night"))

Time_of_day <- na.omit(Time_of_day)
Time_of_day

observed <- mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Evening"]) - mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Night"])
observed

N <- 10^4-1

sample.size = nrow(Time_of_day)
group.1.size = nrow(Time_of_day[Time_of_day$time_of_day == "Evening",])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Time_of_day$dep_delay[index]) - median(Time_of_day$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```

Permutation test between Morning and Night:

```{r}
Time_of_day <- United_Airlines %>% select(dep_time, dep_delay) %>%
  mutate(time_of_day = case_when(
    dep_time >= 600 & dep_time < 1200 ~ "Morning",
    dep_time >= 1200 & dep_time < 1800 ~ "Afternoon",
    dep_time >= 1800 & dep_time < 2100 ~ "Evening",
    TRUE ~ "Night"))

Time_of_day <- na.omit(Time_of_day)
Time_of_day

observed <- mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Morning"]) - mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Night"])
observed

N <- 10^4-1

sample.size = nrow(Time_of_day)
group.1.size = nrow(Time_of_day[Time_of_day$time_of_day == "Morning",])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Time_of_day$dep_delay[index]) - median(Time_of_day$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```

Comparing by dividing the day into two categories:

Permutation test between Morning and Night:

```{r}
Time_of_day <- United_Airlines %>% select(dep_time, dep_delay) %>%
  mutate(time_of_day = case_when(
    dep_time >= 600 & dep_time < 1800 ~ "Morning",
    TRUE ~ "Night"))

Time_of_day <- na.omit(Time_of_day)
Time_of_day

observed <- mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Morning"]) - mean(Time_of_day$dep_delay[Time_of_day$time_of_day == "Night"])
observed

N <- 10^4-1

sample.size = nrow(Time_of_day)
group.1.size = nrow(Time_of_day[Time_of_day$time_of_day == "Morning",])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Time_of_day$dep_delay[index]) - median(Time_of_day$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```


2. Time of year

```{r}
Time_of_year <- United_Airlines %>% select(month, dep_delay) %>%
  mutate(month = factor(month, labels = month.abb))

Time_of_year
ggplot(Time_of_year, aes(x = month, y = dep_delay)) +
  geom_boxplot(color = "cadetblue") +
  labs(x = "Time of Year",
       y = "Departure Delay",
       title = "Time of Year versus Departure Delays for United Airlines (UA)",
       )
summary(Time_of_year)
```

Permutation Test:

```{r}
Time_of_year <- United_Airlines %>% select(month, dep_delay) %>%
  mutate(month = case_when( month >5 & month < 10 ~ "Summer",
                            month <=5 & month >= 10 ~ "Other seasons"))

Time_of_year <- na.omit(Time_of_year)
Time_of_year

observed <- mean(Time_of_year$dep_delay[Time_of_year$month == "Summer"]) - mean(Time_of_year$dep_delay[Time_of_year$month != "Other seasons"])
observed

N <- 10^4-1

sample.size = nrow(Time_of_year)
group.1.size = nrow(Time_of_year[Time_of_year$month == "Summer",])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Time_of_year$dep_delay[index]) - median(Time_of_year$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```


3. Temperature

```{r}
Temperature <- United_Airlines %>% select(temp, dep_delay)
  
Temperature
ggplot(Temperature, aes(x = temp, y = dep_delay)) +
  geom_point(color = "cadetblue") +
  labs(
    x = "Temperature",
    y = "Departure Delay",
    title = "Temperature versus Departure Delays for United Airlines (UA)"
  )
summary(Temperature)
```

Permutation Test between lower temperature and higher temperature:

```{r}
Temperature <- United_Airlines %>% select(temp, dep_delay)

Temperature <- na.omit(Temperature)
Temperature

observed <- mean(Temperature$dep_delay[Temperature$temp <= 57.92]) - mean(Temperature$dep_delay[Temperature$temp > 57.92])
observed

N <- 10^4-1

sample.size = nrow(Temperature)
group.1.size = nrow(Temperature[Temperature$temp <= 57.92,])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Temperature$dep_delay[index]) - median(Temperature$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```


4. Wind Speed

```{r}
Wind_Speed <- United_Airlines %>% select(wind_speed, dep_delay)
  
Wind_Speed
ggplot(Wind_Speed, aes(x = wind_speed, y = dep_delay)) +
  geom_point(color = "cadetblue") +
  labs(
    x = "Wind Speed",
    y = "Departure Delay",
    title = "Wind Speed versus Departure Delays for United Airlines (UA)"
  )
summary(Wind_Speed)
```

Permutation Test between lower wind speed and higher wind speed:

```{r}
Wind_Speed <- United_Airlines %>% select(wind_speed, dep_delay)

Wind_Speed <- na.omit(Wind_Speed)
Wind_Speed

observed <- mean(Wind_Speed$dep_delay[Wind_Speed$wind_speed <= 9.206]) - mean(Wind_Speed$dep_delay[Wind_Speed$wind_speed > 9.206])
observed

N <- 10^4-1

sample.size = nrow(Wind_Speed)
group.1.size = nrow(Wind_Speed[Wind_Speed$wind_speed <= 9.206,])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Wind_Speed$dep_delay[index]) - median(Wind_Speed$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```


5. Precipitation

```{r}
Precipitation <- United_Airlines %>% select(precip, dep_delay)

Precipitation
ggplot(Precipitation, aes(x = precip, y = dep_delay)) +
  geom_point(color = "cadetblue") +
  labs(x = "Precipitation",
       y = "Departure Delay",
       title = "Precipitation versus Departure Delays for United Airlines (UA)",
       )
summary(Precipitation)
```

Permutation test between zero and non-zero Precipitation:

```{r}
Precipitation <- United_Airlines %>% select(precip, dep_delay)

Precipitation <- na.omit(Precipitation)
Precipitation

observed <- mean(Precipitation$dep_delay[Precipitation$precip <= 0]) - mean(Precipitation$dep_delay[Precipitation$precip > 0])
observed

N <- 10^4-1

sample.size = nrow(Precipitation)
group.1.size = nrow(Precipitation[Precipitation$precip <= 0,])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Precipitation$dep_delay[index]) - median(Precipitation$dep_delay[-index])
}

p_value <- 2*(sum(result <= observed) + 1) / (N + 1)
p_value
```



6. Visibility

```{r}
Visibility <- United_Airlines %>% select(visib, dep_delay)

Visibility
ggplot(Visibility, aes(x = visib, y = dep_delay)) +
  geom_point(color = "cadetblue") +
  labs(x = "Visibility",
       y = "Departure Delay",
       title = "Visibility versus Departure Delays for United Airlines (UA)",
       )
summary(Visibility)
```

Permutation Test between <10 visibility and >10 visibility:

```{r}
Visibility <- United_Airlines %>% select(visib, dep_delay)

Visibility <- na.omit(Visibility)
Visibility

observed <- mean(Visibility$dep_delay[Visibility$visib <= 9.251]) - mean(Visibility$dep_delay[Visibility$visib > 9.251])
observed

N <- 10^4-1

sample.size = nrow(Visibility)
group.1.size = nrow(Visibility[Visibility$visib <= 9.251,])
result <- numeric(N)

for (i in 1:N)
  {index = sample(sample.size, size=group.1.size, replace = FALSE)
  result[i] = median(Visibility$dep_delay[index]) - median(Visibility$dep_delay[-index])
}

p_value <- 2*(sum(result >= observed) + 1) / (N + 1)
p_value
```

